Continuous risk-based predictive approaches to treatment effect heterogeneity: A simulation study

### 2021-06-23 submission/manuscript.rmd
update figure captions

```{r rmsebase, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_base.png"))
```

@@

```{r rmsesamplesize, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 17,000; \\textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_sample_size.png"))
```

@@

```{r rmseauc, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.85 and sample size of 4,250; \\textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_auc.png"))
```

The constant effects model, the linear interaction model and models with RCS
smoothing (3 knots) had the highest median c-for-benefit in the base case
scenario and the scenarios where linear and quadratic deviations were
considered. However, constant treatment effects and linear interaction models
tended to present much lower variability compared to all other model-based and
smoothing approaches (Figure \ref{fig:discrimination}). We also observed an
increasing trend of discrimination for benefit variability and decreasing median
values with increasing number of restricted cubic spline knots in all scenarios.

```{r discrimination, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

@@

```{r calibration, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/calibration_base.png"))
```

>>>>

```{r rmsebase, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated superpopulation of size 500,000. \\textit{Panel A} presents the results of the scenario with true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \\textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_base.png"))
```

@@

```{r rmsesamplesize, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 17,000; \\textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \\textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_sample_size.png"))
```

@@

```{r rmseauc, echo=FALSE, fig.cap="RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.85 and sample size of 4,250; \\textit{Panel B} presents the results under moderate linear deviations from constant treatment effects, while holding true prediction AUC and sample size constant; \\textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/rmse_auc.png"))
```

The constant effects model, the linear interaction model and models with RCS
smoothing (3 knots) had the highest median c-for-benefit in the base case
scenario and the scenarios where linear and quadratic deviations were
considered. The constant treatment effect model and the linear interaction model
tended to present much lower variability compared to all other approaches
(Figure \ref{fig:discrimination}). We also observed an increasing trend of
discrimination for benefit variability and decreasing median values with
increasing number of restricted cubic spline knots in all scenarios.

```{r discrimination, echo=FALSE, fig.cap="Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \\textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/discrimination_base.png"))
```

@@

```{r calibration, echo=FALSE, fig.cap="Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \\textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \\textit{Panel B} presents the results under moderate linear deviations from constant treatment effectst; \\textit{Panel C} presents the results for strong linear deviations from of constant relative treatment effects; \\textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.", fig.show="hold", out.width = '100%'}
knitr::include_graphics(here::here("figures/calibration_base.png"))
```

### 2021-06-23 submission/manuscript.tex
update figure captions

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_base} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.}\label{fig:rmsebase}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_sample_size} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 17,000; \textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.}\label{fig:rmsesamplesize}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_auc} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.85 and sample size of 4,250; \textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.}\label{fig:rmseauc}
\end{figure}

The constant effects model, the linear interaction model and models with
RCS smoothing (3 knots) had the highest median c-for-benefit in the base
case scenario and the scenarios where linear and quadratic deviations
were considered. However, constant treatment effects and linear
interaction models tended to present much lower variability compared to
all other model-based and smoothing approaches (Figure
\ref{fig:discrimination}). We also observed an increasing trend of
discrimination for benefit variability and decreasing median values with
increasing number of restricted cubic spline knots in all scenarios.

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/discrimination_base} \caption{Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.}\label{fig:discrimination}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/calibration_base} \caption{Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from the base case of constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from the base case scenario of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from the base case scenario of constant relative treatment effects.}\label{fig:calibration}
\end{figure}

>>>>

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_base} \caption{RMSE of the considered methods across 500 replications calculated in a simulated superpopulation of size 500,000. \textit{Panel A} presents the results of the scenario with true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.}\label{fig:rmsebase}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_sample_size} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 17,000; \textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations constant relative treatment effects.}\label{fig:rmsesamplesize}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/rmse_auc} \caption{RMSE of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.85 and sample size of 4,250; \textit{Panel B} presents the results under moderate linear deviations from constant treatment effects, while holding true prediction AUC and sample size constant; \textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.}\label{fig:rmseauc}
\end{figure}

The constant effects model, the linear interaction model and models with
RCS smoothing (3 knots) had the highest median c-for-benefit in the base
case scenario and the scenarios where linear and quadratic deviations
were considered. The constant treatment effect model and the linear
interaction model tended to present much lower variability compared to
all other approaches (Figure \ref{fig:discrimination}). We also observed
an increasing trend of discrimination for benefit variability and
decreasing median values with increasing number of restricted cubic
spline knots in all scenarios.

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/discrimination_base} \caption{Discrimination for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from constant treatment effects; \textit{Panel C} presents the results for strong linear deviations from constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.}\label{fig:discrimination}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{/home/arekkas/Documents/Projects/arekkas_HteSimulation_XXXX_2021/figures/calibration_base} \caption{Calibration for benefit of the considered methods across 500 replications calculated in a simulated sample of size 500,000. \textit{Panel A} presents the results under the base case scenario of true constant relative treatment effect, with a true prediction AUC of 0.75 and sample size of 4250; \textit{Panel B} presents the results under moderate linear deviations from constant treatment effectst; \textit{Panel C} presents the results for strong linear deviations from of constant relative treatment effects; \textit{Panel D} presents the results for strong quadratic deviations from constant relative treatment effects.}\label{fig:calibration}
\end{figure}
